---
title: "PHC 6937 (Data Visualization) Final Project"
author: "Jack Leary"
date: "`r Sys.Date()`"
output:
  html_document:
    theme: journal
    highlight: tango
    df_print: kable
    toc: true
    toc_float: true
    code_folding: show
    code_download: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      message = FALSE, 
                      warning = FALSE, 
                      fig.align = "center")
options(future.globals.maxSize = 2000 * 1024^2)
```

# Introduction

# Libaries

```{r}
library(dplyr)
library(Seurat)
library(ggplot2)
library(SCISSORS)
library(patchwork)
library(paletteer)
library(kableExtra)
```

# Data

There are several samples here, so we'll need to create a dictionary of the samplenames and the sample types. 

```{r}
dir_names <- c("GSM4679532", "GSM4679533", "GSM4679534", "GSM4679535", "GSM4679536", 
               "GSM4679537", "GSM4679538", "GSM4679539", "GSM4679540", "GSM4679541", 
               "GSM4679542", "GSM4679543", "GSM4679544", "GSM4679545", "GSM4679546", "GSM4679547")
sample_names <- c("P01", "P02", "P03", "P04", "P05", "P06", "P07", "P08", "P09", "P10", 
                  "MET01", "MET02", "MET03", "MET04", "MET05", "MET06")
sample_types <- c(rep("Primary", 10), rep("Metastatic", 6))
sample_dict <- data.frame(Directory = dir_names, 
                          Sample = sample_names, 
                          Type = sample_types)
sample_dict %>% 
  kbl(booktabs = TRUE, caption = "Sample Dictionary") %>% 
  kable_minimal("hover", html_font = "Avenir", full_width = FALSE)
```

# Replicating Lin *et al*

## Dataset Integration

First, we loop across the samples, create `Seurat` objects, and normalize expression and identify variable genes. The first 10 samples are primary tumors, and the next 6 are metastatic. 

```{r}
obj_list <- list()
for (i in seq(sample_dict$Directory)) {
  file <- paste0("~/Desktop/Data/Lin et al/GSE154778_RAW/", sample_dict$Directory[i])
  counts <- Read10X(file)
  seurat_obj <- CreateSeuratObject(counts, 
                                   project = sample_dict$Sample[i], 
                                   min.cells = 3, 
                                   min.features = 1000)
  seurat_obj@meta.data$sample <- sample_dict$Sample[i]
  seurat_obj <- NormalizeData(object = seurat_obj, verbose = FALSE)
  seurat_obj <- FindVariableFeatures(object = seurat_obj, nfeatures = 2000, verbose = FALSE)
  obj_list[[i]] <- seurat_obj
}
primary_obj <- obj_list[1:10]
meta_obj <- obj_list[11:16]
```

```{r, echo=FALSE}
rm(counts, obj_list, seurat_obj, dir_names, file, sample_names, sample_types)
```

Next we integrate all the samples into one `Seurat` object. We'll start with the primary samples. 

```{r}
integration_genes <- SelectIntegrationFeatures(object.list = primary_obj, verbose = FALSE)
integration_anchors <- FindIntegrationAnchors(object.list = primary_obj, 
                                              k.filter = 50, 
                                              anchor.features = integration_genes, 
                                              verbose = FALSE)
primary_orig <- IntegrateData(anchorset = integration_anchors, 
                              k.weight = 50, 
                              verbose = FALSE)
DefaultAssay(primary_orig) <- "integrated"
```

And we repeat the process for the metastatic samples. 

```{r}
integration_genes <- SelectIntegrationFeatures(object.list = meta_obj, verbose = FALSE)
integration_anchors <- FindIntegrationAnchors(object.list = meta_obj, 
                                              k.filter = 50, 
                                              anchor.features = integration_genes, 
                                              verbose = FALSE)
meta_orig <- IntegrateData(anchorset = integration_anchors, 
                           k.weight = 50, 
                           verbose = FALSE)
DefaultAssay(meta_orig) <- "integrated"
```

## Preprocessing

### Primary Tumors

```{r}
primary_orig <- ScaleData(primary_orig, verbose = FALSE)
primary_orig <- RunPCA(primary_orig, 
                       npcs = 30, 
                       verbose = FALSE, 
                       seed.use = 629)
primary_orig <- RunUMAP(primary_orig, 
                        reduction = "pca", 
                        dims = 1:30, 
                        n.components = 2, 
                        seed.use = 629, 
                        verbose = FALSE)
primary_orig <- FindNeighbors(primary_orig, 
                              reduction = "pca", 
                              dims = 1:30, 
                              verbose = FALSE)
primary_orig <- FindClusters(primary_orig, 
                             resolution = 0.3, 
                             verbose = FALSE, 
                             random.seed = 629)
```

### Metastatic Tumors

Now we repeat the preprocessing on the metastatic tumors. 

```{r}
meta_orig <- ScaleData(meta_orig, verbose = FALSE)
meta_orig <- RunPCA(meta_orig, 
                    npcs = 30, 
                    verbose = FALSE, 
                    seed.use = 629)
meta_orig <- RunUMAP(meta_orig, 
                     reduction = "pca", 
                     dims = 1:30, 
                     n.components = 2, 
                     seed.use = 629, 
                     verbose = FALSE)
meta_orig <- FindNeighbors(meta_orig, 
                           reduction = "pca", 
                           dims = 1:30, 
                           verbose = FALSE)
meta_orig <- FindClusters(meta_orig, 
                          resolution = 0.3, 
                          verbose = FALSE, 
                          random.seed = 629)
```

### Visualization

```{r}
p0 <- DimPlot(primary_orig) + 
      labs(x = "UMAP 1", y = "UMAP 2", title = "Primary Tumors") + 
      scale_color_paletteer_d("miscpalettes::pastel") + 
      theme_yehlab() + 
      theme(axis.title = element_text(hjust = 0, size = 10), 
            plot.title = element_text(face = "bold", hjust = 0.5, size = 12)) + 
      guides(color = guide_legend(nrow = 2, override.aes = list(size = 3)))
p1 <- DimPlot(meta_orig) + 
      labs(title = "Metastatic Tumors") + 
      scale_color_paletteer_d("miscpalettes::pastel") + 
      theme_yehlab() + 
      theme(axis.title = element_blank(), 
            plot.title = element_text(face = "bold", hjust = 0.5, size = 12)) + 
      guides(color = guide_legend(nrow = 2, override.aes = list(size = 3)))
(p0 | p1) + plot_annotation(caption = "Louvain Clustering", 
                            theme = theme(plot.caption = element_text(face = "italic")))
```


```{r}
p2 <- DimPlot(primary_orig, group.by = "sample") + 
      labs(x = "UMAP 1", y = "UMAP 2", title = "Sample IDs (Primary Tumors)") + 
      scale_color_paletteer_d("miscpalettes::pastel") + 
      theme_yehlab() + 
      theme(axis.title = element_text(hjust = 0, size = 10), 
            plot.title = element_text(face = "bold", hjust = 0.5, size = 12)) + 
      guides(color = guide_legend(nrow = 2, override.aes = list(size = 3)))
p3 <- DimPlot(meta_orig, group.by = "sample") + 
      labs(title = "Sample IDs (Metastatic Tumors)") + 
      scale_color_paletteer_d("miscpalettes::pastel") + 
      theme_yehlab() + 
      theme(axis.title = element_blank(), 
            plot.title = element_text(face = "bold", hjust = 0.5, size = 12)) + 
      guides(color = guide_legend(nrow = 2, override.aes = list(size = 3)))
(p2 | p3) + plot_annotation(caption = "Louvain Clustering", 
                            theme = theme(plot.caption = element_text(face = "italic")))
```









# New Methodology

